---
title: "Homework 8"
author: "Ahantya Sharma"
date: "2025-04-10"
output: pdf_document
---

### UT EID: as236366

### Github Link: <https://github.com/Ahantya/SDS315/blob/main/HW8/HW8Markdown.Rmd>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(ggplot2)
library(knitr)
library(scales)
library(tidyverse)
library(kableExtra)
library(mosaic)
creatine <- read.csv('creatinine.csv')
covid <- read.csv("covid.csv")
milk <- read.csv("milk.csv")
```

# Problem 1 - regression warm up 

## A.

```{r echo=FALSE, warning=FALSE}

lm0 = lm(creatclear ~ age, data=creatine)


predicted <- predict(lm0, newdata = data.frame(age = 55))

# 113.723
# a. 

#b. is -0.6198


fourtyPredict <- predict(lm0, newdata = data.frame(age = 40)) # predicted to be 123, but 140 HEALTHIER
sixtyPredict <- predict(lm0, newdata = data.frame(age = 60)) # predicted to be 110.62 but 112
```

The creatinine clearance rate we should expect for a 55-year-old is approximately `r round(predicted, 4)`. This was done by finding the linear model of the relationship between a patient's age (in years) and a patient's creatine clearance rate (in mL/minute). Then we used that formula with the intercept as 147.8129 and the slope as -0.6198 and plugged in x (the age) as 55, to find the predicted creatinine clearance rate.

## B. 

The creatinine clearance rate changes with age by -0.6198 mL/minutes per year. This is found by the slope of our linear model (formally named as the age coefficient in the linear model).

## C. 

The creatinine clearance rate is higher (higher) for a 40-year-old with a rate of 135 compared to a 60-year-old with a rate of 112. This is because through the linear model equation (147.8219 - 0.168x), a 60-year-old is expected to have a creatinine clearance rate of `r round(sixtyPredict, 4)` while a 40-year-old is expected to have a creatinine clearance rate of `r round(fourtyPredict, 4)`. Since the difference between the 40-year-old's creatinine clearance rate of 135 to its predicted value is higher compared to the difference between the 60-year-old's creatinine clearance rate of 112 to its predicted value, this means that the 40-year-old's creatinine clearance rate is higher and healthier. 

\newpage


# Problem 2 - Modeling disease growth

## 1.

```{r echo=FALSE}

italy <- subset(covid, country == "Italy")

doublingTime70 <- numeric(10000)

bootItaly <- do(10000) * {
  italyModel <- lm(log(deaths) ~ days_since_first_death, data = resample(italy))
}

bootItalyDoubling <- do(10000) * {
  model <- lm(log(deaths) ~ days_since_first_death, data = resample(italy))
  slope <- coef(model)[2]
  data.frame(doubling = 70 / (slope * 100))
}

ciItaly <- confint(bootItaly, level = 0.95)

kable(ciItaly[ciItaly$name == "days_since_first_death", ], caption = "95% Bootstrapped CI for Growth Rate (Italy)")



kable(confint(bootItalyDoubling, level = 0.95), caption = "95% Bootstrapped CI for Doubling Time (Italy)")



spain <- subset(covid, country == "Spain")


bootSpain <- do(10000) * {
  spainModel <- lm(log(deaths) ~ days_since_first_death, data = resample(spain))
}

```

Through a 95% bootstrapped confidence interval, the estimated growth rate for Italy is about 0.183 with the interval range specifically being about (0.159, 0.209). 

Through a 95% bootstrapped confidence interval, the estimated doubling time for Italy is about 3.8 with the interval range specifically being about (3.4, 4.4). 

## 2. 

```{r echo=FALSE}

bootSpainDoubling <- do(10000) * {
  model <- lm(log(deaths) ~ days_since_first_death, data = resample(spain))
  slope <- coef(model)[2] # gets the slope
  data.frame(doubling = 70 / (slope * 100))
}


ciSpain <- confint(bootSpain, level = 0.95)

kable(
  ciSpain[ciSpain$name == "days_since_first_death", ],
  caption = "95% Bootstrapped CI for Growth Rate (Spain)"
)


kable(
  confint(bootSpainDoubling, level = 0.95),
  caption = "95% Bootstrapped CI for Doubling Time (Spain)"
)
```

Through a 95% bootstrapped confidence interval, the estimated growth rate for Spain is about 0.276 with the interval range specifically being about (0.236, 0.318). 

Through a 95% bootstrapped confidence interval, the estimated doubling time for Spain is about 2.5 with the interval range specifically being about (2.2, 3.0). 

## 3. 

```{r echo=FALSE}

ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line() +  # Line graph
  labs(
    title = "Reported Daily COVID-19 Deaths Over Time",
    x = "Days Since First Death",
    y = "Number of Daily Deaths",
    color = "Country"
  ) +
  scale_color_manual(values = c("Italy" = "red", "Spain" = "blue")) + 
  theme_minimal() 

```

As the line graph shows, Spain's number of daily deaths increased at a faster slope than Italy's number of daily deaths since the first death (up to the point of around 26 days since the first death). 

\newpage

# Problem 3 - price elasticity of demand 

```{r echo=FALSE}


bootMilk <- do(10000) * {
  milkModel <- lm(log(sales) ~ log(price), data = resample(milk))
}

ciMilk <- confint(bootMilk, level = 0.95)

kable(
  ciMilk[ciMilk$name == "log.price.", ],
  caption = "95% Bootstrapped CI for Price Elasticity of Demand (Milk)"
)


# -1.77 to -1.45



```

I estimated the price elasticity of demand by fitting a log-log linear model of sales on price, where the coefficient of log(price) gives the elasticity. Through a bootstrapped 95% confidence interval that finds the linear model, the estimated price elasticity of demand for milk is about -1.619, with the interval range specifically being about (-1.770, -1.456). This means that a 1% increase in price leads to approximately a 1.619 decrease in the quantity demanded. 


