---
title: "Homework 8"
author: "Ahantya Sharma"
date: "2025-04-10"
output: pdf_document
---

### UT EID: as236366

### Github Link: <https://github.com/Ahantya/SDS315/blob/main/HW8/HW8Markdown.Rmd>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
library(ggplot2)
library(knitr)
library(scales)
library(tidyverse)
library(kableExtra)
library(mosaic)
creatine <- read.csv('creatinine.csv')
```

# Problem 1 - regression warm up 

## A.

```{r echo=FALSE, warning=FALSE}

lm0 = lm(creatclear ~ age, data=creatine)


predicted <- predict(lm0, newdata = data.frame(age = 55))

# 113.723
# a. 

#b. is -0.6198


fourtyPredict <- predict(lm0, newdata = data.frame(age = 40)) # predicted to be 123, but 140 HEALTHIER
sixtyPredict <- predict(lm0, newdata = data.frame(age = 60)) # predicted to be 110.62 but 112
```

The creatinine clearance rate we should expect for a 55-year-old is approximately `r round(predicted, 4)`. This was done by finding the linear model of the relationship between a patient's age (in years) and a patient's creatine clearance rate (in mL/minute). Then we used that formula with the intercept as 147.8129 and the slope as -0.6198 and plugged in x (the age) as 55, to find the predicted creatinine clearance rate.

## B. 

The creatinine clearance rate changes with age by -0.6198 mL/minutes per year. This is found by the slope of our linear model (formally named as the age coefficient in the linear model).

## C. 

The creatinine clearance rate is higher (higher) for a 40-year-old with a rate of 135 compared to a 60-year-old with a rate of 112. This is because 